Tail Indices and Scale Parameters in Financial Time Series

### 2017-02-02 papers/FX/draft.tex
Some advancement.
1. Power utility function with gamma < 1 2. some progress on finding the right eigenfunction in the paper with Jeff.

\input{../../physics_common}
\title{Tail Indices and Scale Parameters in Financial Time Series}
\author{Casper de Vries, Thomas Valentin Mikosch, Xie Xiaolei}
\date{\today}
\begin{document}
\maketitle

@@

Rewriting \eqref{eq:xxie0} gives
\begin{eqnarray*}
  u(v) + b u(\delta v) &=& \E u(C) + b \E [u(C); C < \delta v] \\
  &=& \alpha p M^\alpha{1 + b \over 1 - \gamma}\int_{-\infty}^{-M}
  [(1 - \phi) e^r + \phi e^x]^{1 - \gamma} |x|^{-\alpha-1} dx \\
  && + {1 + b \over 1 - \gamma}\int_{-M}^{\delta v}
  [(1 - \phi) e^r + \phi e^x]^{1 - \gamma} d F_u(x) \\
  && + \int_{\delta v}^{\infty}
  [(1 - \phi) e^r + \phi e^x]^{1 - \gamma} d F_u(x) \\
  &=& I_1 + I_2 + I_3
\end{eqnarray*}
Only the first term $I_1$ involves the tail parameters $\alpha$
and $M$ in which we are interested, so we will only compute this
term. The integral of $I_1$ can be written as
\begin{eqnarray*}
&& \int_{-\infty}^{-M}
[(1 - \phi) e^r + \phi e^x]^{1 - \gamma} |x|^{-\alpha-1} dx \\
&=& \int_{M}^{\infty}
[(1 - \phi) e^r + \phi e^{-x}]^{1 - \gamma} x^{-\alpha-1} dx
\end{eqnarray*}
Taylor expansion at $\phi e^{-x} = 0$ gives
\begin{eqnarray*}
  && [(1 - \phi) e^r + \phi e^{-x}]^{1 - \gamma} \\
  &=& ((1 - \phi) e^r)^{1 - \gamma}
  + \sum_{n=1}^{\infty} {1 \over n!}
  {\Gamma(2 - \gamma) \over \Gamma(2 - n -\gamma)}
  ((1 - \phi) e^r)^{1 - n - \gamma} \phi^n e^{-n x}
\end{eqnarray*}
Therefore
\begin{eqnarray*}
&& \int_{-\infty}^{-M}
[(1 - \phi) e^r + \phi e^x]^{1 - \gamma} |x|^{-\alpha-1} dx \\
&=&
{1 \over \alpha} M^{-\alpha} ((1 - \phi) e^r)^{1 - \gamma}
+ \sum_{n=1}^{\infty} {1 \over n!}
{\Gamma(2 - \gamma) \over \Gamma(2 - n -\gamma)}
((1 - \phi) e^r)^{1 - n - \gamma} n^\alpha \Gamma(-\alpha, n M)
\end{eqnarray*}
where $\Gamma(\alpha, x)$ is the upper incomplete gamma
function. For $x > 1$, this function may be expressed as (equation
8.11.2 of {\it NIST Digital Library of Mathematical Functions}
\cite{NIST.DLMF})
\[
\Gamma(a, z) = z^{a - 1} e^{-z}
\sum_{k=0}^\infty (-1)^k {\Gamma(a) \over \Gamma(a - k)} z^{-k}
\]
Thus after re-arranging terms we get
\begin{eqnarray}
  {1 \over 1 + b} I_3
  &=&
  {p \over 1 - \gamma} ((1 - \phi) e^r)^{1 - \gamma} \nonumber \\
  &&
  + \sum_{k=0}^\infty
  {\alpha p ((1 - \phi) e^r)^{1 - \gamma}\over M^{k+1}}
  \prod_{i=0}^{k-1} (\alpha + i)
  \sum_{n=1}^\infty {(-1)^{n-1} \over n! n^{k+1}}
  \left(
  {\phi \over e^{M + r} (1 - \phi)}
  \right)^n \times \nonumber \\
  && \prod_{i=1}^{n-1} (\gamma + i - 1)
  \label{eq:xxie1}
\end{eqnarray}
Observe that the sum over $n$ does not involve $\alpha$. Denote
\[
a_n(k) =
{(-1)^{n-1} \over n! n^{k+1}}
\left(
     {\phi \over e^{M + r} (1 - \phi)}
     \right)^n
     \prod_{i=1}^{n-1} (\gamma + i - 1)
\]
For each given $k$, $a_n(k)$ is an alternating series.
\begin{eqnarray*}
  && {a_n(k) \over |a_{n+1}(k)|} \\
  &=& {
    e^{M + r} (1 - \phi) (n+1)^{k+2}
    \over
    \phi n^{k+1} (n + \kappa - 1)
  } \\
  &=&
  \left(1 + {2 - \kappa \over n + \kappa - 1}\right)
  \left( 1 + {1 \over n} \right)^{k+1}
  {e^{M + r} (1 - \phi) \over \phi}
\end{eqnarray*}

If ${e^{M + r} (1 - \phi) \over \phi} > 1$, it is clear
$a_n(k)/|a_{n+1}(k)| > 1$. It follows that $I_3$ and hence $u(v) +
u(\delta v)$ is monotonically increasing with $\alpha$. If all
investors are guided by and hence seek to maximize the same utility
function, they would all 
have constructed their portfolios by choosing from a set of equities
that have tail indices close to the largest in the market. At market
equilibrium, all actively traded equities would have tail indices that
are indistinguishable from each other.

@@

\subsubsection{very risk-avert utility functions}
In this section we consider power utility functions of
\eqref{eq:power_utility} with $\gamma > 1$. For convenience, let $ĸ =
\gamma - 1$.

>>>>

\input{../../physics_common}
\title{Tail Indices and Scale Parameters in Financial Time Series}
\author{Thomas Valentin Mikosch, Casper de Vries, Xie Xiaolei}
\date{\today}
\begin{document}
\maketitle

@@

Rewriting \eqref{eq:xxie0} gives
\begin{eqnarray*}
  u(v) + b u(\delta v) &=& \E u(C) + b \E [u(C); C < \delta v] \\
  &=& \alpha p M^\alpha{1 + b \over 1 - \gamma}\int_{-\infty}^{-M}
  [(1 - \phi) e^r + \phi e^x]^{1 - \gamma} |x|^{-\alpha-1} dx \\
  && + {1 + b \over 1 - \gamma}\int_{-M}^{\delta v}
  [(1 - \phi) e^r + \phi e^x]^{1 - \gamma} d F_u(x) \\
  && + \int_{\delta v}^{\infty}
  [(1 - \phi) e^r + \phi e^x]^{1 - \gamma} d F_u(x) \\
  &=& I_1 + I_2 + I_3
\end{eqnarray*}
Only the first term $I_1$ involves the tail parameters $\alpha$
and $M$ in which we are interested, so we will only compute this
term. The integral of $I_1$ can be written as
\begin{eqnarray*}
&& \int_{-\infty}^{-M}
[(1 - \phi) e^r + \phi e^x]^{1 - \gamma} |x|^{-\alpha-1} dx \\
&=& \int_{M}^{\infty}
[(1 - \phi) e^r + \phi e^{-x}]^{1 - \gamma} x^{-\alpha-1} dx
\end{eqnarray*}
Taylor expansion at $\phi e^{-x} = 0$ gives
\begin{eqnarray*}
  && [(1 - \phi) e^r + \phi e^{-x}]^{1 - \gamma} \\
  &=& ((1 - \phi) e^r)^{1 - \gamma}
  + \sum_{n=1}^{\infty} {1 \over n!}
  {\Gamma(2 - \gamma) \over \Gamma(2 - n -\gamma)}
  ((1 - \phi) e^r)^{1 - n - \gamma} \phi^n e^{-n x}
\end{eqnarray*}
Therefore
\begin{eqnarray*}
&& \int_{-\infty}^{-M}
[(1 - \phi) e^r + \phi e^x]^{1 - \gamma} |x|^{-\alpha-1} dx \\
&=&
{1 \over \alpha} M^{-\alpha} ((1 - \phi) e^r)^{1 - \gamma}
+ \sum_{n=1}^{\infty} {1 \over n!}
{\Gamma(2 - \gamma) \over \Gamma(2 - n -\gamma)}
((1 - \phi) e^r)^{1 - n - \gamma} n^\alpha \Gamma(-\alpha, n M)
\end{eqnarray*}
where $\Gamma(\alpha, x)$ is the upper incomplete gamma
function. For $x > 1$, this function may be expressed as (equation
8.11.2 of {\it NIST Digital Library of Mathematical Functions}
\cite{NIST.DLMF})
\[
\Gamma(a, z) = z^{a - 1} e^{-z}
\sum_{k=0}^\infty (-1)^k {\Gamma(a) \over \Gamma(a - k)} z^{-k}
\]
Thus after re-arranging terms we get
\begin{eqnarray}
  {1 \over 1 + b} I_1
  &=&
  {p \over 1 - \gamma} ((1 - \phi) e^r)^{1 - \gamma} \nonumber \\
  &&
  + \sum_{k=0}^\infty
  {\alpha p ((1 - \phi) e^r)^{1 - \gamma}\over M^{k+1}}
  \left[ \prod_{i=0}^{k-1} (\alpha + i) \right]
  \sum_{n=1}^\infty {(-1)^{n-1} \over n! n^{k+1}}
  \left(
  {\phi \over e^{M + r} (1 - \phi)}
  \right)^n \times \nonumber \\
  && \prod_{i=1}^{n-1} (\gamma + i - 1)
  \label{eq:xxie1}
\end{eqnarray}
Observe that the sum over $n$ does not involve $\alpha$. Denote
\[
a_n(k) =
{(-1)^{n-1} \over n! n^{k+1}}
\left(
     {\phi \over e^{M + r} (1 - \phi)}
     \right)^n
     \prod_{i=1}^{n-1} (\gamma + i - 1)
\]
For each given $k$, $a_n(k)$ is an alternating series.
For $n = 1,3,5,\cdots$
\begin{eqnarray*}
  && {a_n(k) \over |a_{n+1}(k)|} \\
  &=& {
    e^{M + r} (1 - \phi) (n+1)^{k+2}
    \over
    \phi n^{k+1} (n + \kappa - 1)
  } \\
  &=&
  \left(1 + {2 - \kappa \over n + \kappa - 1}\right)
  \left( 1 + {1 \over n} \right)^{k+1}
  {e^{M + r} (1 - \phi) \over \phi}
\end{eqnarray*}

If ${e^{M + r} (1 - \phi) \over \phi} > 1$, it is clear
$a_n(k)/|a_{n+1}(k)| > 1$. It follows that $I_1$ and hence $u(v) +
u(\delta v)$ is monotonically increasing with $\alpha$. If all
investors are guided by and hence seek to maximize the same utility
function, they would all 
have constructed their portfolios by choosing from a set of equities
that have tail indices close to the largest in the market. At market
equilibrium, all actively traded equities would have tail indices that
are indistinguishable from each other by standard estimation procedures.

Furthermore, differentiating \eqref{eq:xxie1} gives
\begin{eqnarray*}
  && \opd{\phi} [u(v) + b u(\delta v)] \\
  &=&
  \pd{I_1}{\phi} + \opd{\phi}(I_2 + I_3) \\
  &=&
  -(1 + b) {
    p ((1 - \phi) e^r)^{1- \kappa}
    \over
    1 - \phi
  }
  - (1 + b) \sum_{k=0}^\infty \sum_{n=1}^\infty
  {\alpha p \over M^{k + 1}}
  {\Gamma(\alpha + k) \over \Gamma(\alpha)}
  {(-1)^{n-1} \over n! n^{k+1}} \times \\
  &&
  {\Gamma(\gamma + n - 1) \over \Gamma(\gamma)}
  {1 \over \phi}
  \left[
    {\phi \over e^{M+r} (1 - \phi)}
  \right]^n
  e^r [(1 - \gamma) \phi + n] [(1 - \phi) e^r]^{-\gamma} \\
  && + \partial_\phi I_2 + \partial_\phi I_3
\end{eqnarray*}
where we have used the short-hand notation $\partial_\phi I_i =
\pd{I_i}{\phi}$. Truncating the infinite sum above after $k=0$ and
$n=1$, and then setting the result to 0 provides a condition that the
optimal choice of $\phi$ must satisfy:
\begin{eqnarray*}
  {
    p [(1 - \phi) e^r]^{-\gamma}
    [e^{-M} \alpha (1 - \phi (1 - \gamma)) - e^r M (1 - \phi)]
    (1 + b)
    \over
    M (1 - \phi)
  } + \partial_\phi I_2 + \partial_\phi I_3 = 0
\end{eqnarray*}
Solving for $\alpha$, we get
\begin{eqnarray*}
  \alpha &=& {
    M e^M (1 - \bar \phi)
    \left[
      e^r p (1 + b) - ((1 - \bar \phi) e^r)^\gamma
      (\partial_\phi I_2 + \partial_\phi I_3)_{\phi = \bar \phi}
    \right]
    \over
    p [1 - (1 - \gamma) \bar \phi] (1 + b)
  }
\end{eqnarray*}
where $\bar \phi$ denotes the optimal choice of $\phi$, the portion of
one's wealth allocated to the equity.

@@

\subsubsection{very risk-avert utility functions}
In this section we consider power utility functions of
\eqref{eq:power_utility} with $\gamma > 1$. For convenience, let $\kappa =
\gamma - 1$.
